---
title: "Florida Aviation Elevation Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Overview

This project evaluates the elevation distribution of aviation facilities in Florida and maps both the mean center and the elevation-weighted mean center of airport locations.

## Load Data

```{r}
aviation <- read.csv("aviationfacilities.csv", header = TRUE)
str(aviation)
```

## Filter Florida Facilities

```{r}
aviation_fl <- subset(aviation, STATE_CODE == "FL")
```

## Descriptive Statistics

```{r}
summary(aviation_fl$ELEV)
```

```{r}
mean_fl   <- mean(aviation_fl$ELEV, na.rm = TRUE)
median_fl <- median(aviation_fl$ELEV, na.rm = TRUE)
sd_fl     <- sd(aviation_fl$ELEV, na.rm = TRUE)
range_fl  <- range(aviation_fl$ELEV, na.rm = TRUE)
iqr_fl    <- IQR(aviation_fl$ELEV, na.rm = TRUE)

mean_fl
median_fl
sd_fl
range_fl
iqr_fl
```

## Key Statistical Findings

For aviation facilities in Florida, the mean elevation is 69.1 m and the median elevation is 55.0 m. Elevation ranges from 0.0 m to 709.0 m, with an interquartile range of 78.6 m and a standard deviation of 64.2 m.

```{r}
boxplot(
  aviation_fl$ELEV,
  main = "Elevation of Aviation Facilities in Florida",
  ylab = "Elevation (m)"
)
```

## Interpretation

The boxplot shows multiple extreme observations, so the median provides a more robust summary of central tendency than the mean.

## Mean Center and Weighted Mean Center

```{r}
FL_airports <- aviation_fl[
  aviation_fl$SITE_TYPE_CODE == "A", ]

plot(
  FL_airports$LONG_DECIMAL,
  FL_airports$LAT_DECIMAL,
  xlab = "Longitude",
  ylab = "Latitude",
  main = "Airports in Florida",
  xlim = c(-88, -80),
  ylim = c(24, 33)
)

n <- nrow(FL_airports)[1]

mc_x <- sum(FL_airports$LONG_DECIMAL) / n
mc_y <- sum(FL_airports$LAT_DECIMAL) / n

points(mc_x, mc_y, pch = 15, cex = 2, col = "blue")

wmc_x <- sum(as.numeric(FL_airports$ELEV * FL_airports$LONG_DECIMAL), na.rm = TRUE) /
         sum(FL_airports$ELEV, na.rm = TRUE)

wmc_y <- sum(as.numeric(FL_airports$ELEV * FL_airports$LAT_DECIMAL), na.rm = TRUE) /
         sum(FL_airports$ELEV, na.rm = TRUE)

points(wmc_x, wmc_y, pch = 15, cex = 2, col = "red")

legend(
  "topright",
  legend = c("Mean centre", "Weighted mean centre"),
  pch = c(15, 15),
  col = c("blue", "red")
)
```
